The development of abstract concepts in children's early lexical networks

### 2019-02-28 CMCL/ms.tex
ms

\title{How to Make a Camera-Ready Proceedings Contribution}

@@

\begin{abstract}
This document contains the instructions for preparing a camera-ready
manuscript for the proceedings of EACL-2017. The document itself
conforms to its own specifications, and is therefore an example of what
your manuscript should look like. These instructions should be used for
both papers submitted for review and for final versions of accepted
papers. Authors are asked to conform to all the directions reported in
this document.
\end{abstract}

\section{Methods}\label{methods}

\subsection{Nouns}\label{nouns}

We used data from Wordbank (Frank et al., 2017), an open repository
aggregating cross-linguistic language developmental data of the
MacArthur-Bates Communicative Development Inventory (CDI), a parent
report vocabulary checklist, Toddler version. Parent report is a
reliable and valid measure of children's vocabulary (Fenson et al.,
1994). We used a set of nouns representing hand-checked translation
equivalents across 10 languages (available in Wordbank database). We
``assume'' (i should ask what were the criteria for the unilemma
selection) that this set of translation equivalents is representative of
a core shared vocabulary across all children (in different cultures) by
around 30 months,

The conceptual organization derived from this shared vocabulary is
obviousely identical across languages, but we investigate whether there
are systematic cross-lingusitc variations in the order of aquisition of
words that make up this vocabulary, and crucially, whether such
variation influences the induced conceptual organization at different
points in time. In fact, differences in the order of acqusition of words
do not necessarily give rise to different conceptual organization.
Imagine that two languages vary in whether ``cow'' or ``dog'' is
acquired first. This difference will not change the induced conceptual
organization across time since both ``dog'' and ``cow'' are instances of
the same high-level concept ``animal''.

\begin{CodeChunk}
\captionsetup{width=0.8\textwidth}\begin{figure}[h]

{\centering \includegraphics{figs/unnamed-chunk-1-1}

@@

\caption[bla bla]{bla bla}\label{fig:unnamed-chunk-1}
\end{figure}
\end{CodeChunk}

\subsection{Features}\label{features}

We used two sources of information which represent two ways words may be
related to one another in the semantic network. The first source was the
MacRae features (McRae et al., 2005). These features were collected by
giving adult participants a set of nouns and prompting them to provide
various kinds of properties (perceptual, functional, encyclopedic, and
taxonomic). We followed Hills et al. 2010 in excluding the encyclopedic
and taxonomic properties as their role may not be important at this
stage of development. The other source was the semantic similarity
derived based on co-occurrence in the corpus of CHILDES, using word2vec.

\subsection{Network}\label{network}

We constructed networks using CDI words as nodes. Pairs of words were
linked by an edge weighted by the number of shared MacRae feature or
with the continuous distance obtained via Word2vec. To model change in
the conceptual organization from month to month, we constructed a
different network at each month, based on the words that have been
acquired by that month. We defined the age of acquisiton of a given word
by the month at which this noun was produced by at least 50\% of
children, in each language (Goodman et al., 2008).

\subsection{Small World properties}\label{small-world-properties}

We test whether the networks display the so-called ``small-world''
properties similar to other semantic and real-world networks (Steyvers
\& Tenenbaum, 2005; Watts \& Strogatz, 1998). Small world properties are
characterized with the average clustering coefficient \(C\) and the
average shortest path \(L\). The former measures the extent to which the
network is clustered, i.e., made of highly connectned sub-networks,
whereas the latter measures the typical separation between tow nodes in
the network. A network is small-word if it has a higher clustering
coeffient comapred to a randomly connected network of the same size
\(C \gg C_{random}\), while still having a shorest path length as small
as the one typically observed in random networks, that is
\(L \approx L_{random}\).

@@

{\centering \includegraphics{figs/unnamed-chunk-2-1}

@@

\caption[bla bla]{bla bla}\label{fig:unnamed-chunk-2}
\end{figure*}
\end{CodeChunk}

\subsection{Clustering analysis}\label{clustering-analysis}

We identified network clusters using WalkTrap (Pons \& Latapy, 2006), a
community detection algorithm based on the fact that a random walker
tends to be trapped in dense parts of a network corresponding to highly
interconnected sub-groups of nodes.

\subsection{Evaluation of clusterings aross
development}\label{evaluation-of-clusterings-aross-development}

We note \(\mathcal{C}_t\) the clustering at month \(t\), based on the
subset of the vocabulary acquired by that month in each language. The
evaluation consists in comparing \(\mathcal{C}_t\) to the final
clustering \(\mathcal{C}^*\) obtained using the full vocabulary by the
last month of acquisition (i.e., \(\mathcal{C}^*\)=\(\mathcal{C}_t\) for
\(t\)=30 months). Such comparison allows us to study how words acquired
at different points in time may induce changes in the conceptual
organization. In addition, we investigate consistency and variability in
this change across languages.

We compare \(\mathcal{C}_t\) to \(\mathcal{C}^*\) using a standard
method in clustering comparison, which is based on counting word pairs
on which the two clusterings agree or disagree (Rand, 1971; Hubert and
Arabie, 1985). A pair of words learned by month \(t\) can fall under one
of the four follwoing cases: 1) pairs that are placed in the same
cluster under \(\mathcal{C}_t\) and in the same cluster under
\(\mathcal{C}^*\) (True positives, noted as \(tp(\mathcal{C}_t)\)), 2)
pairs placed in different clsuters under \(\mathcal{C}_t\) and in
different clusters under \(\mathcal{C}^*\) (True negatives,
\(tn(\mathcal{C}_t)\)), 3) pairs placed in the same cluster under
\(\mathcal{C}_t\) and in different clusters under \(\mathcal{C}^*\)
(False positive, \(fp(\mathcal{C}_t)\)), and 4) pairs placed in
different clusters under \(\mathcal{C}_t\) and in the same cluster under
\(\mathcal{C}^*\) (False negatives, \(fn(\mathcal{C}_t)\)).

The clustering precison \(P(\mathcal{C}_t)\) and recall
\(R(\mathcal{C}_t)\) are defiend as follows:

\[
P(\mathcal{C}_t) = \frac{|tp(\mathcal{C}_t)|}{|tp(\mathcal{C}_t)| + |fp(\mathcal{C}_t)|}
\] \[
R(\mathcal{C}_t) = \frac{|tp(\mathcal{C}_t)|}{|tp(\mathcal{C}_t)| + |fn(\mathcal{C}_t)|}
\] Both Precison and Recall converge to 1 (perfect score) as
\(\mathcal{C}_t\) becomes more and more similar to \(\mathcal{C}^*\). If
precision starts low before converging to 1 (as opposed to being a
constant at 1), this pattern would indicate that some pairs that should
be differentiated are initially lumped together, suggesting a process of
``differentiation'' over development. Similarly, if we observe an
increase in Recall, this pattern would indicate that some pairs that
should be associated are initially differentiate, suggesting a process
of ``coelescence'' over development.

\subsection{Word ordering}\label{word-ordering}

The question we are addressing is whether the order of acqusition of
words affect the

@@

For real and random: results show that both precision and recall
increase across development, suggesting that the conceptual organization
undergoes both differentiation for some word pairs (example), and
coelescecne for other pairs (examples).

For within-cluster: precision is high, but recall is low: explain why

How to explain this change?

\begin{quote}
Assuming a random/real ranking of words: -Some words may be lamped
together initially (even if we force a relatively high number of
clusters) because of the smaller vocabulary size? Can we have an example
of this? Are the other clsuters empty? -Some words may be differentiated
initially because of the forcing? -Is there a roleof noise? maybe
clustering is just more random for smaller size vocabulary?
\end{quote}

\begin{quote}
Assuming a with-cluster ranking of words: -Some words may be lamped
together initially (even if we force a relatively high number of
clusters): this does not happen, at least while we are still working
with the first cluster\ldots{}that's why precision is generally high
-Some words may be differentiated initially because they
\end{quote}

\subsection{Order of word learning}\label{order-of-word-learning}

\section{Sudy of the conceptual
development}\label{sudy-of-the-conceptual-development}

Models

\section{Important citation stuff!
READ!}\label{important-citation-stuff-read}

\subsection{Why we can't have nice
things}\label{why-we-cant-have-nice-things}

So because the ACL committee wants their \texttt{.tex} files all nice
and consistent, the old version of this package isn't good for what they
want. The Markdown-to-Latex conversion would automatically add in the
references and citations, but would literally hardcode them into the
\texttt{.tex} file. We can't have that, now can we? The only ways you
were allowed to cite in your \texttt{.tex} file was with
\texttt{\textbackslash{}cite\{Gusfield97\}},
\texttt{\textbackslash{}shortcite\{Gusfield97\}}, or
\texttt{\textbackslash{}newcite\{Gusfield97\}}, which corresponded to
``(Gusfield, 1997)'', ``(1997)'', or ``Gusfield (1997)'', respectively.

This means you couldn't do anything like ``(e.g., Gusfield, 2017)'', and
their \texttt{.bst} file was incompatible with \texttt{natbib}.
Obviously. However, I have come to the rescue. In the
\texttt{eacl2017.sty} file in this package, I have added my own bit of
magic: the \texttt{\textbackslash{}barecite\{\}} command, which
corresponds to ``Burchill, 2017''. Notice this version doesn't have
parentheses, so you can get stuff like ``(e.g., Burchill, 2017)'' with
\texttt{(e.g.,\ \textbackslash{}barecite\{Gusfield97\})}. Just be
careful about making sure you don't forget parentheses. Also not that
when you submit the \texttt{.zip} folder, you should make sure to
include the edited \texttt{eacl2017.sty} file.

\subsection{Examples}\label{examples}

\begin{itemize}
\tightlist
\item
  \texttt{\textbackslash{}cite\{Gusfield97\}} becomes \cite{Gusfield97}
\item
  \texttt{\textbackslash{}shortcite\{Gusfield97\}} becomes
  \shortcite{Gusfield97}
\item
  \texttt{\textbackslash{}newcite\{Gusfield97\}} becomes
  \newcite{Gusfield97}
\item
  \texttt{\textbackslash{}barecite\{Gusfield97\}} becomes
  \barecite{Gusfield97}
\item
  \texttt{\textbackslash{}cite\{Gusfield97,Aho72\}} becomes
  \cite{Gusfield97,Aho72}
\item
  \texttt{\textbackslash{}barecite\{Gusfield97,Aho72\}} becomes
  \barecite{Gusfield97,Aho72}
\end{itemize}

\textbf{Wait, \emph{are those citations appearing as question marks?}
Don't worry, read on!}

\subsection{The two-step process of knitting this
file}\label{the-two-step-process-of-knitting-this-file}

Because the ACL's you-need-to-use-our-citing-function formatting doesn't
work well with R Markdown--which really likes to compile its own
citations--knitting this is now a two-step process. After you get it how
you want it in RMarkdown, knit as usual. You'll see that all citations
are question marks, and that the bibliography is missing. This is
natural, don't panic.

For the current purposes, let's assume this file we're in right now is
called \texttt{acl\_draft.Rmd}. Navigate into the directory containing
\texttt{acl\_draft.Rmd} via a terminal. Then run the following command
(swapping out your actual file name for ``acl\_draft'')

\texttt{pdflatex\ acl\_draft.tex;\ pdflatex\ acl\_draft.tex;\ bibtex\ acl\_draft.aux;\ pdflatex\ acl\_draft.tex;\ pdflatex\ acl\_draft.tex}

Now, check the \texttt{.pdf} file. If all went well, it should now have
all the citations in it as well as the reference section. Good luck!

\section{General Formatting
Instructions}\label{general-formatting-instructions}

All the default content below is lifted directly from Kyle MacDonald's
Cogsci 2016 project (\url{https://github.com/kemacdonald/cogsci2016}),
which this entire package is based on. I haven't changed it, so it's up
to you to ignore as you please.

For general information about authoring in markdown, see
**\url{http://rmarkdown.rstudio.com/authoring_basics.html.**}

For standard spoken papers and standard posters, the entire contribution
(including figures, references, everything) can be no longer than six
pages. For abstract posters, the entire contribution can be no longer
than one page. For symposia, the entire contribution can be no longer
than two pages.

The text of the paper should be formatted in two columns with an overall
width of 7 inches (17.8 cm) and length of 9.25 inches (23.5 cm), with
0.25 inches between the columns. Leave two line spaces between the last
author listed and the text of the paper. The left margin should be 0.75
inches and the top margin should be 1 inch.
\textbf{The right and bottom margins will depend on whether you use
U.S. letter or A4 paper, so you must be sure to measure the width of
the printed text.} Use 10 point Times Roman with 12 point vertical
spacing, unless otherwise specified.

The title should be in 14 point, bold, and centered. The title should be
formatted with initial caps (the first letter of content words
capitalized and the rest lower case). Each author's name should appear
on a separate line, 11 point bold, and centered, with the author's email
address in parentheses. Under each author's name list the author's
affiliation and postal address in ordinary 10 point type.

Indent the first line of each paragraph by 1/8\textasciitilde{}inch
(except for the first paragraph of a new section). Do not add extra
vertical space between paragraphs.

\section{First-Level Headings}\label{first-level-headings}

First level headings should be in 12 point , initial caps, bold and
centered. Leave one line space above the heading and
1/4\textasciitilde{}line space below the heading.

\subsection{Second-Level Headings}\label{second-level-headings}

Second level headings should be 11 point , initial caps, bold, and flush
left. Leave one line space above the heading and 1/4\textasciitilde{}
line space below the heading.

\subsubsection{Third-Level Headings}\label{third-level-headings}

Third-level headings should be 10 point , initial caps, bold, and flush
left. Leave one line space above the heading, but no space after the
heading.

\section{Formalities, Footnotes, and
Floats}\label{formalities-footnotes-and-floats}

\subsection{Footnotes}\label{footnotes}

Indicate footnotes with a number\footnote{Sample of the first
footnote.} in the text. Place the footnotes in 9 point type at the
bottom of the page on which they appear. Precede the footnote with a
horizontal rule.\footnote{Sample of the second footnote.}

\subsection{Figures}\label{figures}

All artwork must be very dark for purposes of reproduction and should
not be hand drawn. Number figures sequentially, placing the figure
number and caption, in 10 point, after the figure with one line space
above the caption and one line space below it. If necessary, leave extra
white space at the bottom of the page to avoid splitting the figure and
figure caption. You may float figures to the top or bottom of a column,
or set wide figures across both columns.

\subsection{Two-column images}\label{two-column-images}

You can read local images using png package for example and plot it like
a regular plot using grid.raster from the grid package. With this method
you have full control of the size of your image. \textbf{Note: Image
must be in .png file format for the readPNG function to work.}

You might want to display a wide figure across both columns. To do this,
you can change the \texttt{fig.env} chunk option to \texttt{figure*}. To
align the image in the center of the page, set \texttt{fig.align} option
to \texttt{center}.

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/2-col-image-1}

\caption[This image spans both columns]{This image spans both columns.}\label{fig:2-col-image}
\end{figure*}
\end{CodeChunk}

\subsection{One-column images}\label{one-column-images}

Single column is the default option, but if you want set it explicitly,
set \texttt{fig.env} to \texttt{figure}.

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/image-1}

\caption[One column image]{One column image.}\label{fig:image}
\end{figure}
\end{CodeChunk}

\subsection{R Plots}\label{r-plots}

You can use R chunks directly to plot graphs. And you can use latex
floats in the fig.pos chunk option to have more control over the
location of your plot on the page. For more information on latex
placement specifiers see
**\url{https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions**}

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/plot-1}

\caption[R plot]{R plot}\label{fig:plot}
\end{figure}
\end{CodeChunk}

\subsection{Tables}\label{tables}

Number tables consecutively; place the table number and title (in 10
point) above the table with one line space above the caption and one
line space below it, as in Table 1. You may float tables to the top or
bottom of a column, set wide tables across both columns.

You can use the xtable function in the xtable package.

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
(Intercept) & 0.07 & 0.09 & 0.7 & 0.49 \\ 
  x & 1.92 & 0.09 & 20.4 & 0.00 \\ 
   \hline
\end{tabular}
\caption{This table prints across one column.} 
\end{table}

\section{Acknowledgements}\label{acknowledgements}

Place acknowledgments (including funding information) in a section at
the end of the paper.

>>>>

\title{The development of abstract concepts in children's early lexical
networks}

@@

\begin{abstract}
How do children learn abstract concepts such as animal vs.~artifact?
Previous research has suggested that such concepts can partly be derived
using cues from the language children hear around them. Following this
suggestion, we propose a model where we represent the children's
developing lexicon as an evolving network. The nodes of this network are
based on vocabulary knowledge as reported by parents, and the edges
between pairs of nodes are based on the probability of their
co-occurrence in a corpus of child-directed speech. We found that
several abstract categories can be identified as the dense regions in
such networks. In addition, our simulations suggest that these
categories develop simultaneously, rather than sequentially, thanks to
the children's word learning trajectory which favors the exploration of
the global conceptual space.
\end{abstract}

\section{Introduction:}\label{introduction}

One of the central challenges in cognitive development is to understand
how concepts develop \cite{carey2009,keil1992,gopnik1997}. Of particular
interest is the case of abstract concepts which have non-obvious shared
properties such as ``animal'' and ``artifact''. For example, a cat and a
bird are perceptually quite different but they share some fundamental
properties (e.g., breathing, feeding, and reproducing) which make them
animals (as opposed to artifacts). In such cases, learning requires in
part cultural/linguistic cues which provide information beyond what can
be obtained through the senses \cite{gelman2009,harris2012,csibra2009}.

One way children's conceptual learning can benefit from the language
they hear around them is through word co-occurrence. For example, one
can learn an abstract concept (e.g., animal) simply by observing how its
instances (e.g., ``cat'' and ``bird'') go together in speech. Indeed,
previous work has shown that the caregiver's input contains rich
co-occurrence information about various abstract concepts
\cite{huebner2018}. This work, however, has explored the conceptual
space from an adult perspective (using the words uttered by the
caregivers). Here we explore how abstract concepts may develop from the
children's perspective, investigating how their word learning trajectory
influences the higher-level organization of their developing lexicon.

We study development in light of two hypothetical models. On the first,
called the simultaneous model, learning starts by exploring all parts of
the conceptual space; categories are refined simultaneously over
development. On the second, called the sequential model, learning starts
by exploring a small region of the conceptual space (e.g., the category
``animals'') and only after the refinement of this category, does the
learner move to another.

The paper is organized as follows. First, we describe the research
strategy. In brief, we represented the developing lexicon as an evolving
network and we used word co-occurrence in parent speech as a measure of
words' relatedness. We operationalized abstract concepts as the highly
interconnected regions of the network. Second, we explore how the
pattern of children's word learning influences higher-level conceptual
development, and whether this development corresponds to a simultaneous
or a sequential conceptual growth.

{\centering \includegraphics{figs/network-1}

@@

\caption[Network obtained using a sample of nouns in CDI data (nodes), and co-occurrence-based similarity from a corpus of child-directed speech (edges)]{Network obtained using a sample of nouns in CDI data (nodes), and co-occurrence-based similarity from a corpus of child-directed speech (edges). Colors indicate highly interconnected clusters identified using unsupervised network community detection. The clusters correspond, overall, to four higher-level concepts: animal, food, clothes, and artifacts.}\label{fig:network}
\end{figure*}
\end{CodeChunk}

\section{Data and Methods}\label{data-and-methods}

\subsection{Constructing lexical
networks}\label{constructing-lexical-networks}

The networks' nodes were nouns from Wordbank \cite{frank2017}, an open
repository aggregating cross-linguistic developmental data of the
MacArthur-Bates Communicative Development Inventory (CDI), a parent
report vocabulary checklist, Toddler version \cite{fenson94}. Pairs of
nouns were linked by weighted edges representing their semantic
similarity derived based on co-occurrence in the corpus of
child-directed speech CHILDES \cite{macwhinney2014}, using Word2Vec
algorithm \cite{mikolov2013}.

First, we constructed the end-state network based on all nouns learned
by the last age of acquisition. We used a subset of CDI nouns for which
cross-linguistic translations are present, allowing us to explore
cross-linguistic variability. We used data from the ten following
languages: Croatian, Danish, English, French, Italian, Norwegian,
Russian, Spanish, Swedish, and Turkish. The size of this subset varied
from 314 in Russian (representing 100\% of total nouns in this language)
to 176 in Turkish (representing 59.26\% of total nouns). Second, in
order to study development towards the end-state, we constructed a
different network at each month, based on the nouns that have been
learned by that month.

\subsection{Identifying abstract concepts in a
network}\label{identifying-abstract-concepts-in-a-network}

We assume that abstract concepts correspond to clusters of highly
interconnected nodes in the networks. We identified such clusters using
WalkTrap \cite{pons2006}, an unsupervised community detection algorithm
based on the fact that a random walker tends to be trapped in dense
parts of a network. Figure \ref{fig:network} shows the outcome of
cluster identification in the end-state network in English. The
algorithm obtained four major clusters corresponding to the categories
of clothes, food, animal and artifacts. We refer to this end-state
clustering as \(\mathcal{C}^*\). To examine developmental change in the
conceptual organization, we ran the cluster identification algorithm at
each month of acquisition \(t\), and we compared the resulting
clustering, noted \(\mathcal{C}_t\), to that of the end-state
\(\mathcal{C}^*\). The method of this comparison is detailed below.

@@

{\centering \includegraphics{figs/results-1}

@@

\caption[Mean precision and recall scores obtained through comparing the end-state clustering to clusterings at different months of acquisition, averaged across languages and numbers of clusters]{Mean precision and recall scores obtained through comparing the end-state clustering to clusterings at different months of acquisition, averaged across languages and numbers of clusters. Colors indicates real and hypothetical word sampling mechanisms. Errors bars represent 95\% confidence intervals.}\label{fig:results}
\end{figure*}
\end{CodeChunk}

\subsection{Measuring conceptual
development}\label{measuring-conceptual-development}

We measure conceptual development by comparing \(\mathcal{C}_t\) to
\(\mathcal{C}^*\) across time. We used a standard method in clustering
comparison, which is based on word pairs on which the two clusterings
agree or disagree \cite{rand1971,hubert1985}. We quantify clustering
comparison using precision \(P(\mathcal{C}_t)\) and recall
\(R(\mathcal{C}_t)\), defined as follows:

\[
P(\mathcal{C}_t) = \frac{|tp(\mathcal{C}_t)|}{|tp(\mathcal{C}_t)| + |fp(\mathcal{C}_t)|}
\]

\[
R(\mathcal{C}_t) = \frac{|tp(\mathcal{C}_t)|}{|tp(\mathcal{C}_t)| + |fn(\mathcal{C}_t)|}
\] Where \(tp(\mathcal{C}_t)\) are the true positives, defined as the
word pairs that are placed in the same cluster under \(\mathcal{C}_t\)
and in the same cluster under \(\mathcal{C}^*\). \(fp(\mathcal{C}_t)\)
are the false positives defined as the pairs placed in the same cluster
under \(\mathcal{C}_t\) and in different clusters under
\(\mathcal{C}^*\). Finally, \(fn(\mathcal{C}_t)\) are the false
positives defined as the pairs placed in different clusters under
\(\mathcal{C}_t\) and in the same cluster under \(\mathcal{C}^*\).

We made this comparison using different degrees of clustering
granularity. More precisely, we fixed the same number of clusters for
\emph{both} \(\mathcal{C}_t\) and \(\mathcal{C}^*\), and we varied this
number from two to four clusters. We did not use the trivial case of one
cluster, nor did we did use more than four clusters, since this number
was optimal for the largest network (i.e., the end-state network) based
on the modularity maximization criterion \cite{newman2006}.

\subsection{Learning mechanisms}\label{learning-mechanisms}

We examined how abstract concepts develop under an average word learning
trajectory derived from real developmental data. To construct this
trajectory, we used the normative age of acquisition, that is, the age
at which a word is produced by at least 50\% of children in each
language \cite{goodman2008}. We compared this development to the
development induced by a simultaneous model and the development induced
by a sequential model.

The simultaneous model was instantiated as a uniform sampling across
time from the end-state vocabulary. The sequential model had the
additional constraint of sampling from one category at a time: the first
word is selected randomly from one cluster, subsequent words are sampled
from the same cluster. After all words from this cluster are used, a
word from a different cluster is chosen, and the same process is
repeated until all clusters are covered.

@@

Figure \ref{fig:results} shows the scores obtained through comparing
\(\mathcal{C}^*\) to \(\mathcal{C}_t\) at different points in time
\(t\). For the real word learning trajectory, both precision and recall
start relatively low, indicating that the induced conceptual
organization is initially quite different from that of the end-state.
Both measures converge towards 1 (i.e., perfect score) as
\(\mathcal{C}_t\) becomes more and more similar to \(\mathcal{C}^*\).

The simultaneous model mimics closely the patterns of real conceptual
development, explaining almost all the variance in mean precision
(\(R^2 =\) 0.94) and recall (\(R^2 =\) 0.99). In contrast, the
sequential model had generally a higher precision, i.e., it induced
fewer false positive pairs. This result is due to the fact that we
sampled instances from the same category. However, the same model had
generally lower recall scores, i.e., it induced more false negative
pairs. This second result was due to the fact that sampling from the
same category leads to clusterings that are finer in their conceptual
granularity than the end-state. As a consequence of this discrepancy
with respect to real development, the sequential model explained less
variance than the simultaneous model did in both its mean precision
(\(R^2 =\) 0.44) and recall (\(R^2 =\) 0.96).

\section{Discussion}\label{discussion}

Can children learn abstract concepts based on word co-occurrence in the
language they hear around them? Previous work has shown that
child-directed speech contains information about several abstract
concepts \cite{huebner2018}. Here we investigated when and how this
information becomes available to children as their lexical network
grows. We found that even with a small lexicon, several high-level
concepts such as ``animal'', ``artifact'', ``food'' and ``clothes''
emerge bottom-up as clusters of highly interconnected nodes in the
network. Furthermore, compared with a model that posited sequential
learning, we found that these categories tended to emerge in concert
with one another.

The development of the higher-level conceptual structure seems to be
unaffected by the order with which words are acquired (as long as this
order approximates a uniform sampling from the end-state lexicon),
suggesting that the process of conceptual development can accommodate a
wide range of word learning trajectories without a qualitative change in
the higher-level organization. For example, whether acquisition starts
first with the words ``cat'' and ``banana'' or with the words ``cow''
and ``potato'' does not qualitatively affect the higher-level
organization involving ``animal'' and ``food''. This property is
important as it suggests, for instance, that development is resilient to
variability in the children's linguistic input
\cite{slobin2014,hart1995}.

Developmental changes were captured by precision and recall. The
increase in precision means that false positives decrease over time:
some word pairs that are initially lumped together in a same category,
are eventually differentiated. Similarly, the increase in recall means
that false negatives decrease, that is, some word pairs that are
initially distinct, become eventually subsumed by a same category. These
patterns suggest a process of conceptual reorganization involving both
``differentiation'' and ``coalescence'' as has been suggested in the
developmental literature \cite{carey2009}.

That said, these developmental changes were not necessarily related to
specific concepts (since the patterns were similar in the simultaneous
model where we randomized the order of word learning). Instead, this
finding suggests that differentiation and coalescence of word pairs in
our data are related to the change in the vocabulary size across
development: As more words are added to their lexical network, learners
may approximate better the underlying conceptual organization of the
mature lexicon and would make fewer categorization errors. Indeed,
research in network science indicates that properties of a real network
become more distorted as the size of a sampled sub-network decreases
\cite{leskovec2006}.

One limitation of this study is that we used the normative age of
acquisition, computed using different children at different age groups.
This choice was due to the cross-sectional nature of available CDI data.
Though such a measure has been widely used to study important aspects of
the early lexical networks \cite{hills2009,stella2017,storkel2009}, it
only applies at the population level. In our case, though we found that
concepts develop simultaneously, individual children may display, at
least locally, a sequential-like behavior. For example, prior knowledge
about dinosaurs may enable the learning of new dinosaur-related words
more easily \cite{chi1983}.

In sum, this work provided a quantitative account of how abstract
concepts can emerge from the interaction of the children's emerging
vocabulary and the properties of their linguistic input. One important
direction for future work is to investigate the extent to which the
correlational findings obtained in this study (e.g., the identity of
categories formed across development or the fact that categorization
errors decrease with the size of the lexicon) can be corroborated by
controlled behavioral experiments.

\vspace{1em}

\fbox{\parbox[b][][c]{7.3cm}{\centering All data and code will be available online}}
\vspace{1em}
